---
title: "Statistical inference with the GSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
```

### Load data

```{r load-data}
# Import the dataset
obesity_data <- read.csv("Obesity_Data.csv")

# Identify the character columns
char_cols <- sapply(obesity_data, is.character)

# Convert character columns to factors for easier analysis
obesity_data[char_cols] <- lapply(obesity_data[char_cols], as.factor)
```

* * *

## Part 1: Data

### How are the observations in the sample collected?
The provided dataset encompasses details regarding the classification of individuals' obesity. It has been compiled from diverse sources, including medical records, surveys, and self-reported data.
Due to this being an observational study, this analysis can only be used for generalized purposes; no causal connections can be made.

* * *

## Part 2: Research question
Estimate the true average weight based on this sample using a 95% Confidence Interval.

* * *

## Part 3: Exploratory data analysis

#### Variables being considered
- Weight

### Investigate the Dataset and Variables
```{r eda}
# Weight
summary(obesity_data$Weight)
str(obesity_data$Weight)
hist(obesity_data$Weight)
boxplot(obesity_data$Weight)
scatter.smooth(obesity_data$Weight)
qqnorm(obesity_data$Weight)
qqline(obesity_data$Weight)
```

### Data Manipulation and Cleaning
```{r manip_and_clean}
# Keep columns for analysis
weight_data_trimmed <- obesity_data[, c("Weight")]

# Check rows with N/A's and remove if feasible
(num_rows_with_na <- sum(is.na(weight_data_trimmed)))

# Remove rows with N/As (if necessary)
weight_data_cleaned <- na.omit(weight_data_trimmed)

# Analyze cleaned data
str(weight_data_cleaned)
```
* * *

## Part 4: Inference

### Conditions
Conditions for this confidence interval: 
1. Independence: Sampled observations must be independent. 
‣ Random sample/assignment
‣ If sampling without replacement, n < 10% of population
2. Sample size/skew: n ≥ 30, larger if the population distribution is very skewed.

### Confidence Interval

```{r CI}
# Calculate Mean
mean_value <- mean(weight_data_cleaned)
cat("Mean:", mean_value, "\n")

# Calculate Standard Deviation
sd_value <- sd(weight_data_cleaned)
cat("Standard Deviation:", sd_value, "\n")

# Calculate Standard Error
n <- length(weight_data_cleaned)  # Sample size
standard_error <- sd(weight_data_cleaned) / sqrt(n)
cat("Standard Error:", standard_error, "\n")

## Calculate critical t-score
# Specify the significance level
alpha <- 0.05

# Specify the degrees of freedom
df <- n-1

# Acquire the critical t-score
t_crit <- abs(qt(alpha, df))

# Print the critical t-score
print(t_crit)

## Calculate the confidence interval manually
confidence_interval_min <- mean_value - t_crit * standard_error
confidence_interval_max <- mean_value + t_crit * standard_error
combined <- paste("  CI:", confidence_interval_min, "-", confidence_interval_max)
cat(combined)

## Calculate the confidence interval from the t-test
result <- t.test(weight_data_cleaned)

# Extract confidence interval
confidence_interval <- result$conf.int
cat("Confidence Interval:", confidence_interval, "\n")

# Perform t-test
result <- t.test(weight_data_cleaned, mu = 59)

# Print the test result
print(result)
```

We are 95% confident the average weight is between 54 and 65 Kilograms and the p_value
indicates the true mean is approximately 59.